Exploring the composition of lithic assemblages in Mesolithic south-eastern Norway

### 2021-04-04 analysis/paper/paper.Rmd
Minor edits

Figure \@ref(fig:cor) displays a CA using the lithic count data. While no clear-cut clusters can be readily delineated, the general impression from the plots is still that a chronological dimension is associated with the patterning in the data. This is indicated by the general transition across the colour scale in the row plot, combined with the fact that the two first dimensions of the CA-analysis account for as much as `r round(explained_cumul_dim2, 2)` % of the inertia or variance in the data. The earliest sites tend to be located in the upper right corner of plot A, with increasingly younger sites towards the bottom along the second dimension. The sites from the later parts of the Mesolithic are drawn out along the first dimension of the plot, and are not as impacted by the second dimension.

The column plot reveals that the earliest sites are characterised by the flint artefact categories microburins, projectiles, as well as macro tools and associated debitage. It is also interesting that these sites to larger extent are characerised by core fragments, both in flint and non-flint materials, while the cores themselves appear to be a trait more defining of assemblages with later dates. The non-flint material on the earliest sites appears to be centered around the production of projectiles, as both the projectiles themselves and non-flint blades are important constituents of the assemblages at these sites.

@@

The results of the correspondence analysis does appear to align reasonably well with previous research that is based on more informal analyses of, indicating that the employed artefact categories are behavourally meaningful. Aligning the signficance of this chronological patterning to specific behavioural dimensions, is, however, another task entirely, and a task that is severely hampered by fact that for the most part we lack even a most basic understanding of what any individual lithic object has been used for, and consequently how the different debitage and tool categories should be treated @dibble2017. For example, a vast amount of artefacts defined as debitage are likely to have fulfilled the function of tools, and both debitage and formal tools could have had various different purposes and had a multitude of shapes throughout their use-life. Reducing the number of such inferential leaps by aggregating artefact groups could potentially subsume important variation, but it does also reduce the possibility that the conclusions of any analysis is not simply the result of the employed units of analysis and an overly naive trust in the fact that these units reflect

>>>>

Figure \@ref(fig:cor) displays a CA using the lithic count data. While no clear-cut clusters can be readily delineated, the general impression from the plots is still that a chronological dimension is associated with the patterning in the data. This is indicated by the general transition across the colour scale in the row plot (Figure \@ref(fig:cor)A), combined with the fact that the two first dimensions of the CA-analysis account for as much as `r round(explained_cumul_dim2, 2)` % of the inertia or variance in the data. The earliest sites tend to be located in the upper right corner of plot A, with increasingly younger sites towards the bottom along the second dimension. The sites from the later parts of the Mesolithic are drawn out along the first dimension of the plot, and are not as impacted by the second dimension.

The column plot (Figure \@ref(fig:cor)B) reveals that the earliest sites are characterised by the flint artefact categories microburins, projectiles, as well as macro tools and associated debitage. It is also interesting that these sites to larger extent are characerised by core fragments, both in flint and non-flint materials, while the cores themselves appear to be a trait more defining of assemblages with later dates. The non-flint material on the earliest sites appears to be centered around the production of projectiles, as both the projectiles themselves and non-flint blades are important constituents of the assemblages at these sites.

@@

The results of the correspondence analysis does appear to align reasonably well with previous research that has been based on more informal analyses of artefact types, indicating that the employed artefact categories are capturng some culturally meaningful patterning. Aligning the signficance of these chronological patterns with any specific behavioural dimensions, for example as related to various forms of mobiliy patterns or modes of resource exploitation  is, however, another task entirely. This follows most clearly from the fact that for the most part we lack even a most basic understanding of what any individual lithic object has been used for, and consequently how the different debitage and tool categories should be treated [@dibble2017]. For example, a vast amount of artefacts defined as debitage are likely to have fulfilled the function of tools, and both debitage and formal tool types could have had various different purposes and had a multitude of shapes throughout their use-life. While use-wear analysis can help identify what an artefact was used fo towards the end of its use-file, these are typically only conducted on artefacts already grouped into a single category based on on .  Reducing the number of such inferential leaps by aggregating artefact groups could potentially subsume important variation, but it does also reduce the possibility that the conclusions of any analysis is not simply the result of the employed units of analysis, and an overly naive trust in the fact that these units reflect any behaviourally meaningful categories. This way of viewing lithic assembalges is a fairly established tradition outside Scandinavian Mesolithic research, meaning a range of empirical  measures associated with various mobility patterns are available. Of these the most well-estbalished is the. base don the assumption that, this measure and has been shown to display a meaningful measure across a wide range of temporal, geographical and cultural contexts.
